# Case Studies

## Application Programming Interfaces (API)

An API is nothing else but an interface to facilitate machine to machine 
communication. An interface can be anything, any protocol or pre-defined process.
But of course there are good and bad ways to communicate. Plus some matter-of-taste 
type of decisions. But security and standard compliance are none of the latter. 
There are standards such as the popular, URL based REST that make developers'
lives a lot easier -- regardless of the language they prefer. 

So what do we with an REST API in R or Python? Build a wrapper.
A wrapper is program that wraps an API call in a function using a function's
input parameters to compose the URL that the REST call is. Because, thanks to 
the API's documentation, we know what the output of the API looks like. Hence 
we can process the result and return an R or Python object. 

Here's an example of a very simple API wrapper that makes use of the 
Metropolitan Museum of Modern Art's API to obtain identifiers of pictures based on 
a simple search. 

```{r, eval=FALSE, warning=FALSE,message=FALSE}
# Visit this example query
# https://collectionapi.metmuseum.org/public/collection/v1/search?q=umbrella
# returns a json containing quite a few ids of pictures that were tagged 'umbrella'

#' Search MET
#'
#' This function searches the MET's archive for keywords and
#' returns object ids of search hits. It is a simple wrapper
#' around the MET's Application Programming interface (API).
#' The function is designed to work with other API wrappers
#' and use object ids as an input.
#' @param character search term
#' @return list containing the totoal number of objects found
#' and a vector of object ids.
#'
# Note these declaration are not relevant when code is not
# part of a package, hence you need to call library(jsonlite)
# in order to make this function work if you are not building
# a package.
#' @examples
#' search_met("umbrella")
#' @importFrom jsonlite formJSON
#' @export
search_met <- function(keyword){
  # note how URLencode improves this function
  # because spaces are common in searches
  # but are not allowed in URLs
  url <- sprintf("https://collectionapi.metmuseum.org/public/collection/v1/search?q=%s", URLencode(keyword))
  fromJSON(url)
}

```

You can use these ids with another [endpoint]() in order to receive the pictures
themselves.

```{r, eval=FALSE,warning=FALSE,message=FALSE}
download_met_images_by_id <- function(ids,
                                      download = "primaryImage") {
  # Obtain meta description objects from MET API
  obj_list <- lapply(ids, function(x) {
    req <- download.file(sprintf("https://collectionapi.metmuseum.org/public/collection/v1/objects/%d",
                   x),destfile = "temp.json")
    fromJSON("temp.json")
  })

  # Extract the list elements that contains
  # img URLs in order to pass it to the download function
  img_urls <- lapply(obj_list, "[[", download)
  # Note the implicit return, no return statement needed
  # last un-assigned statement is returned from the function
  lapply(seq_along(img_urls), function(x) {
    download.file(img_urls[[x]],
                  destfile = sprintf("data/image_%d.jpg", x)
    )
  })
}


# Step 4: Use the Wrapper
umbrella_ids <- search_met("umbrella")
umbrella_ids
download_met_images_by_id(umbrella_ids$objectIDs[2:4])
```


Many services such as Google Cloud, AWS, your university library,
your favorite social media account or your local metro operator provide an API. 
Often, in the case of data providers there are even API wrapper packages around for
the most popular programming with data languages. For example the KOF Swiss Economic Institute
provides an API Wrapper package called {kofdata}. The even has an API wrapper to access large parts of their
entire economic database. The R package is called {OECD}. A good way to learn more about creating API wrappers
is to look in to the source code of these packages. 

```{r, warning=FALSE, message=FALSE}
library(kofdata)
# just for viz
library(tstools)
tsl <- get_time_series("ch.kof.barometer")
tsplot(tsl)

```


## Create Your Own API





## Basic Parallel Programming

## The Shiny Web Application Framework

## Spatial Visualization with Leaflet and Open Streetmap




